Week 5 - R Assignment 
========================================================

For this week we are looking at the difference between pre-allocating and appending for various data types.

Vectors
-----------------------------
### Using Numeric Data
#### Appending a 3 digit number to the vector:
```{r}
vector1=function(){
    v=vector()
    for(i in 1:10000){
        v=append(v,123)
    }
    return(v)
}
system.time(vector1())
```
#### Pre-allocating the space before assigning the 3 digit number:
```{r}
vector2=function(){
    v=vector("integer", 10000)
    for(i in 1:10000){
        v[i]=123            
        }
    return(v)
}
system.time(vector2())
```
### Using character data
#### Appending a character string to the vector:
```{r}
vector3=function(){
    v=vector()
    for(i in 1:10000){
        v=append(v,"test")
    }
    return(v)
}
system.time(vector3())
```
#### Pre-allocating the space before assigning the string:
```{r}
vector4=function(){
    v=vector("character", 10000)
    for(i in 1:10000){
        v[i]="test"            
        }
    return(v)
    len
}

system.time(vector4())
```
### Using Factor Data
#### Appending the factor data to a vector:
```{r}
vector5=function(){
    data = c(1)
    fdata = factor(data)
    v=vector()
    for (i in 1:10000){
        v=append(v,fdata)
        }
    return(v)
}
system.time(vector5())
```

#### Pre-allocating the space in the vector first:
```{r}
vector6=function(){
    data = c(1)
    fdata = factor(data)
    v=vector("list", 10000)
    for(i in 1:10000){
        v[i]=fdata            
        }
    return(v)
}
system.time(vector6())
```
#### Conclusion
For all of the above examples we can see that the there is a significant performance hit if we append a value to a vector. It is much more efficient to pre-allocate the space for the vector then to allow the vector to grow by appending data to the end of the list. 

Lists
---------------------------------------------------
### Using Numeric Data
#### Appending a 3 digit number to the list:
```{r}
list1=function(){
    l=list()
    for(i in 1:10000){
        l=append(l,123)
    }
    return(l)
}
system.time(list1())
```
#### Pre-allocating the space before assigning the 3 digit number:
```{r}
list2=function(){
    l=list("integer", 10000)
    for(i in 1:10000){
        l[i]=123            
        }
    return(l)
}
system.time(list2())
```
### Using character data
#### Appending a character string to the list:
```{r}
list3=function(){
    l=list()
    for(i in 1:10000){
        l=append(l,"test")
    }
    return(l)
}
system.time(list3())
```
#### Pre-allocating the space before assigning the string:
```{r}
list4=function(){
    l=list("character", 10000)
    for(i in 1:10000){
        l[i]="test"            
        }
    return(l)
    len
}

system.time(list4())
```
### Using Factor Data
#### Appending the factor data to a list:
```{r}
list5=function(){
    data = c(1)
    fdata = factor(data)
    l=list()
    for (i in 1:10000){
        l=append(l,fdata)
        }
    return(l)
}
system.time(list5())
```

#### Pre-allocating the space in the list first:
```{r}
list6=function(){
    data = c(1)
    fdata = factor(data)
    l=list("list", 10000)
    for(i in 1:10000){
        l[i]=fdata            
        }
    return(l)
}
system.time(list6())
```
#### Conclusion
For all of the above examples we can again see that the there is a significant performance hit if we append a value to a list. It is much more efficient to pre-allocate the space for the list then to  allow the list to grow by appending data to the end of the list. 

Data Frames
-----------------------------------------------------
### Using Numeric Data
#### Appending a 3 digit number to the data frame:
```{r}
df1=function(){
    df <- data.frame(a=numeric(),b=numeric())
    for(i in 1:1000){
        df <- rbind(df, data.frame(a=123,b=456))
    }
    return(df)
}
system.time(df1())
```
#### Pre-allocating the space before assigning the 3 digit number:
```{r}
df2=function(){
    df<-data.frame(a = numeric(1000),b = numeric(1000))
    for(i in 1:1000){
        df$a[i] <- 123
        df$b[i] <- 456
    }
    return(df)
}
system.time(df2())
```
### Using character data
#### Appending a character string to the data frame:
```{r}
df3=function(){
    df <- data.frame(a=character(),b=character(), stringsAsFactors=FALSE)
    for(i in 1:1000){
        df <- rbind(df, data.frame(a="test",b="code"))
    }
    return(df)
}
system.time(df3())
```
#### Pre-allocating the space before assigning the string:
```{r}
df4=function(){
    df<-data.frame(a = character(1000),b = character(1000), stringsAsFactors = FALSE)
    for(i in 1:1000){
        df$a[i] <- "test"
        df$b[i] <- "code"
    }
    return(df)
}
system.time(df4())
```
### Using Factor Data
#### Appending the factor data to data frame:
```{r}
df5=function(){
    df <- data.frame(a=character(),b=character())
    for(i in 1:1000){
        df <- rbind(df, data.frame(a="test",b="code"))
    }
    return(df)
}
system.time(df5())
```
#### Conclusion
Once again for all of the above examples we can again see that the there is a significant performance hit if we append a value to a data frame. It is much more efficient to pre-allocate the space for the datframe first, then to allow the data frame to grow by appending data to the end of the data frame.

The Boston dataset from the MASS Package
=============================================================

### Loading the MASS Package:
```{r}
require(MASS)
```
### Showing that the Boston dataset is loaded:
```{r}
head(Boston)
```
### Creating a new data frame from the Boston dataset and renaming the columns:
```{r}
bostondf <- data.frame(CrimeRate=Boston$crim, EmploymentDistance=Boston$dis, StudentTeacherRatio=Boston$ptratio, MedianHomeValue=Boston$medv)
```
### Showing the head and tail of the new data frame:
```{r}
head(bostondf)
tail(bostondf)
```